---
title: "First Homework"
author: "Jonathan Kuek"
date: "23 September 2015"
output: html_document
---


Q1 Data structures  
In what ways does a list differ from a vector in R? Use R code to create one of each.  

Vectors can only be one of the four types:  

```{r}
char <- c("Homework", "One")
number <- c(2.9, 1.5)
logic <- c(TRUE, FALSE)
int <- c(-1L, 2L)
char
number
logic
int
```

Lists are containers and can contain different classes of elements:  

```{r}
mix <- list(char, number, logic, int)
mix
```

2 Style  
Rewrite the following code so that it is consistent with Hadley Wickham’s R style guide.  

num = 1  
mycondition <- round (runif (1,0,1,)) if( mycondition ) {  
num <- num + 1  
print(paste("num =" , num , sep=" ")) } else {  
print("false") }  

Rewritten code:
```{r}
num <- 1
my_condition <- round(runif(1, 0, 1)) 
if(my_condition) {
  num <- num + 1
  print(paste("num =", num, sep=" ")) 
} else {
  print("false") 
}
```

Q3 Functions  
Write an R function that does the following:  
Inputs a positive numeric vector, verifies it is a positive numeric vector of length greater than one, and returns its geometric mean. You can look up the definition of the geometric mean on Wikipedia and utilize the prod() function.
Utilize the stop() function with a clear error message in the case that the input is invalid. Finally, call your function with some valid input.

```{r}
geometric_mean <- function(input) {
  l <- length(input)
  if(l>1) {
    if(all(input>0)){
      answer <- prod(input)^(1/l)
      print(answer)
    } else {
    stop('Error! Please enter a vector with more than one positive number')
    }
  } else {
    stop('Error! Please enter a vector with more than one positive number')
  }
}

geometric_mean(c(2, 18))
```

Q4 Subsetting a data.frame  
In the RStudio console, call
View(state.x77)
to look at a spreadsheet-like representation of these data.
Then write R code that extracts the data for New York on the variables Murder and Frost in three different ways.
```{r}
state.x77["New York", ]
state.x77["New York", c("Murder", "Frost")]
state.x77["New York", c(5, 7)]
state.x77["New York", c(FALSE, FALSE, FALSE, FALSE, TRUE, FALSE, TRUE)]
```


5 Manipulating data  
In the RStudio console, type  

```{r}
help(airquality)
```
to see information on a data.frame regarding the air quality in New York. Then:
Replace the Month variable with the month represented as an ordered factor.  
```{r}
new_air_quality <- data.frame(airquality)
new_air_quality[1:31, 5] <- "May"
new_air_quality[32:61, 5] <- "June"
new_air_quality[62:92, 5] <- "July"
new_air_quality[93:123, 5] <- "August"
new_air_quality[124:153, 5] <- "September"
new_air_quality
```
